*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-21 Code/Manuscript_O2.Rmd
Edit text in Intro

---
title: "*Prochlorococcus marinus* responses to light and oxygen" 
author:
  - name: '**Mireille Savoie**'
    affil: 1
    main: true
    email: msavoie@mta.ca
  - name: Aurora Mattison
    affil: 1, 2
  - name: Julie Nadeau
    affil: 1
  - name: Laurel Genge
    affil: 1
  - name: Sylwia Śliwińska-Wilczewska
    affil: 1
  - name: Maximilian Berthold
    affil: 1, 3
  - name: Naaman Omar
    affil: 1
  - name: Ondřej Prášil
    affil: 1, 4
    main: false
  - name: Amanda M. Cockshutt
    affil: 1, 5
    main: false
  - name: Douglas A. Campbell
    affil: 1
    main: false
    orcid: '0000-0001-8996-5463'
    email: 'dcampbel@mta.ca'
affiliation:
  - num: 1
    address: Department of Biology, Mount Allison University
institute:
  - mta: 'Department of Biology, Mount Allison University, 63B York St., Sackville NB, Canada, E4L 1G7.'
  - num: 2
  - ubc: 'Department of Biology, University of British Columbia'
  - num: 3
  - umwelt: 'Umweltbundesamtm, Dessau-Roßlau, Germany'
  - num: 4
  - alga: 'Algatech, Czech Academy of Science'
  - num: 5
  - sfx: 'Department of Chemistry, St. Frances Xavier University'
date: "`r format(Sys.Date())`"
output:
  bookdown::word_document2:
    code_folding: show
    reference_docx: KnitTemplate.docx
    keep_md: yes
    toc: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
  bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
always_allow_html: true
bibliography: [Manuscript_O2.bib, RPackages.bib]
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

@@

*Prochlorococcus marinus* comprises many strains, organized into clades, defined by 16S-23S intergenic transcribed ribosomal sequence signatures (Rocap et al., 2003). The clades inhabit distinct ecological niches [@moorePhysiologyMolecularPhylogeny1998], originally defined by Low Light (LL) or High Light (HL).  Of at least 12 known *Prochlorococcus* clades only 5 to date have cultured representatives; HLI, HLII, LLI, LLII/III and LIV [@billerProchlorococcusStructureFunction2015].

@@

Our changing climate, characterized by warming ocean waters, is rapidly altering conditions for these highly specialized strains of marine picophytoplankton. Predictions indicate a net global increase of *P. marinus* cell abundances of 29%, along with  poleward latitudinal shifts of at least 10° in marine phytoplankton niches by the end of this century [@flombaumPresentFutureGlobal2013] in response to warming waters, with increases in *Prochlorococcus* of approximately 50% in the more poleward regions. As the solar incidence angles are lower and photoperiod seasonality is more pronounced at higher latitudes, this shift will carry these phytoplankters into new combinations of photoperiod, seasonal regimes and changing spectral regimes [@bartonAnthropogenicClimateChange2016]. A 10° poleward shift of *P. marinus*, coupled with a dramatic increase in cell abundance between 40°N and 50°N,  could have a large impact on ocean ecosystems and global biogeochemical cycles [@flombaumPresentFutureGlobal2013], and thus will undoubtedly impact the ecosystem balance off our Canadian waters.

Global warming is also rapidly changing the chemistry of our oceans.  By the end of this century, surface ocean pH is projected to decline by 0.1 to 0.4  due to  projected increases in carbon dioxide concentrations [@garcia-sotoOverviewOceanClimate2021]. Moreover, substantial changes in the global water cycle, leading to extensive changes in worldwide precipitation patterns, are affecting ocean salinity levels on a global scale, and  ice melts due to rising temperatures are impacting salinity levels in the Arctic and Northwest Atlantic oceans [@leeRecognizingSalinityThreats2022]. Increasing sea temperatures are also causing a decrease in [O~2~] across our global oceans [@matearLongtermChangesDissolved2003], with significant declines in [O~2~] observed toward the poles in both hemispheres [@helmObservedDecreasesOxygen2011]. Warmer ocean waters increase stratification, and decrease oxygen solubility at the surface, which in turn decreases oxygen mixing downwards by ocean currents [@garcia-sotoOverviewOceanClimate2021]. Models predict that the OMZ in the Pacific and Indian Oceans are expanding [@garcia-sotoOverviewOceanClimate2021; @buseckeDivergingFatesPacific2022]. In light of these expanding OMZ, a recent study by @buseckeDivergingFatesPacific2022 highlights an intriguing finding; the cores of the OMZ, where the oxygen levels are lowest, is projected to contract.  Given the predicted decrease in overall oceanic [O~2~],projected expansion of OMZs, and possible decline in the OMZ core volumes, we set forth to examine the influence of [O~2~] on *P. marinus* growth, as this area of research has been relatively understudied.

@@

We then analyzed growth rates in terms of cumulative diel Photosynthetically Usable Radiation (PUR). Photosynthetic organisms rely on absorbing light energy within the PAR range, 350 to 700 nm, for photosynthesis [@morelAvailableUsableStored1978]. PUR represents the fraction of PAR that is absorbed by the pigments of photosynthetic organisms [@morelAvailableUsableStored1978]. PUR, therefore, takes into account the pigment composition of photosynthetic organisms and the specific spectral wavebands these pigments absorb.  *Prochlorococcus marinus* light-harvesting complexes, [@chisholmProchlorococcusMarinusNov1992;
@goerickePigmentsProchlorococcusMarinus1992], show an absorption maxima
of 442 nm for divinyl chlorophyll a and 478 nm for divinyl chlorophyll b
[@goerickePigmentsProchlorococcusMarinus1992] allowing *P. marinus* to
efficiently harvest blue light in the 400 nm
to 500 nm range [@morelAvailableUsableStored1978] prevailing in deep ocean
habitats, where only blue spectral wavelengths prevail [@holtropVibrationalModesWater2021].  In *Prochlorococcus* small cell diameters, from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992], and simple cell structures, minimize the complication of pigment package effect or intracellular self-shading [@morelProchlorococcusSynechococcusComparative1993]. Given the different spectral light regimes typical of the niches of different ecotypes,  expressing growth rates in terms of cumulative diel absorbed PUR might simplify different photoperiods, spectral bands, and PAR levels into a common parameter, making growth response comparisons across strains and different oxygen levels more accessible.  We also aimed to detect whether growth responses are driven simply by cumulative diel absorbed PUR, or whether specific photoperiods, spectral bands or PAR levels have independent, albeit interacting, effects on growth.

@@

Three xenic cultures of *P. marinus* were obtained from Bigelow Labs, NCMA
Maine, US (REFERENCE). MED4 (CCMP1986) is from High Light-adapted (HLI) clade; SS120
(CCMP1375) is from Low Light-adapted (LLIII) clade; and MIT9313 (CCMP2773)
is from Low Light-adapted (LLIV) clade. Cultures were
maintained in incubators set to 22°C with a light/dark cycle of 12 h. The PAR level for maintenance cultures reflected Photosynthetically Active Radiation (PAR) in the source
niche of the ecotype; MED4, of 160 µmol photons m^-2^ s^-1^ with illumination from STANDARD Products Inc. Cool White F24T5/41K/8/HO/PS/G5/STD, 24 watts, fluorescent bulbs; SS120 and
MIT9313 at 30 µmol photons m^-2^ s^-1^ with illumination from Philips Cool White F14T5/841 Alto, 14 watts, fluorescent bulbs. To maintain active growth all strains were
transferred weekly with 1 to 5 dilution with Pro99 media
[@mooreCulturingMarineCyanobacterium2007] prepared with autoclaved artificial
seawater [@bergesEvolutionArtificialSeawater2001].

@@

The growth rate of MED4, from the high light (HLI) clade of *P.
marinus*, increased with higher imposed PAR and longer photoperiods
(Figure \@ref(fig:MED4PAR)), when grown under 250 µM O~2~. No growth was
observed under any imposed conditions under a 4 h photoperiod. MED4 was
able to grow under the lowest PAR treatment of 30 µmol photons m^-2^
s^-1^ but only when the blue light photoperiod was longer than 4 h; the
white LED light photoperiod was longer than 8 h; or when the red light
photoperiod was longer than 16 h. Growth rate increased
with increasing photoperiods for all spectral wavebands when exposed to
PAR of 90 and 180 µmol photons m^-2^ s^-1^. MED4 grew fastest for each
spectral waveband under 16 h photoperiods. The maximum growth rate
(µ~max~) was
`r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 180 uE blue
light and 16 h photoperiod.

@@

MED4PAR_cap <- glue("**Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MED4 (High Light (HLI) near surface clade) vs. photoperiod (h).  **",

@@

The growth rate of SS120, from the low light (LLIII) clade of *P.
marinus*, increased with lower imposed PAR and longer photoperiods, when
grown under 250 µM O~2~ (Figure \@ref(fig:SS120PAR)). SS120 was able to
grow under the lowest PAR treatment of 30 µmol photons m^-2^ s^-1^ but
only when the photoperiod was longer than 4 h under blue, white, and red
light. SS120 required longer than 8 h photoperiod of imposed green
light. No growth was observed under any blue light photoperiods when
exposed to PAR of 90 µmol photons m^-2^ s^-1^. Growth rate increased
with increasing photoperiods for white, red and green light when exposed
to PAR of 90 µmol photons m^-2^ s^-1^; however, no growth was observed
when the red light photoperiod was 16 h. Experiments conducted under PAR
of 180 µmol photons m^-2^ s^-1^ showed no growth under imposed blue and
red lights. Growth rate decreased with longer photoperiods and showed
growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol
photons m^-2^ s^-1^ and white LED light. The growth rate generally
increased with longer photoperiods under green light treatments, however
no growth was observed under the 8 h photoperiod. The µ~max~ was
`r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^  achieved under 90 µmol photons
m^-2^ s^-1^ white LED light and 16 h photoperiod.

@@

SS120PAR_cap <- glue("**Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* SS120 (Low Light (LLIII) deep ocean clade) vs. photoperiod (h). **",

@@

The growth rate of MIT9313, from the low light (LLIV) clade of *P.
marinus*, increased with low to moderate imposed PAR and longer
photoperiods, when grown under 250 µM O~2~ (Figure
\@ref(fig:MIT9313PAR)). Growth rate increased with increasing
photoperiods under the lowest PAR treatment of 30 µmol photons m^-2^
s^-1^ for blue, white LED and red light. The green light treatment
showed growth only when the photoperiod was longer than 12 h and
decreased by \~ 50% under a 16 h photoperiod. Under PAR of 90 µmol
photons m^-2^ s^-1^, growth rate increased with longer photoperiods for
white LED, red, and green light treatments. Growth rates decreased from
4 to 8 h blue light photoperiod and exhibited no growth under
photoperiods of 12 and 16 h. Blue light did not induce growth at 180
µmol photons m^-2^ s^-1^ under any photoperiod; white LED and red light
treatments only exhibited growth under the 8 h photoperiod and growth
rates increased with longer photoperiods under green light. The µ~max~
was
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 30 µmol photons
m^-2^ s^-1^ blue light and 16 h photoperiod.

@@

MIT9313PAR_cap <- glue("**Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MIT9313 (Low Light (LLIV) deep ocean clade) vs. photoperiod (h). **",

@@

AllStrainPurFitsPlots_cap <- glue("**Chlorophyll specific growth rate (d^-1^) vs. cumulative diel Photosynthetic Usable Radiation (PUR, µmol photons m^-2^ d^-1^).**",
                    "**A.** MED4 High Light I ecotype under 250, 25, or 2.5 µM of O~2~.",
                    "**B.** SS120 Low Light II/III (LLII/III) ecotype under 250, 25, or 2.5 µM of O~2~.",
                    "**C.** MIT9313 Low Light IV (LLIV) ecotype under 250, 25, or 2.5 µM of O~2~.",
                    "3 levels of imposed dissolved O~2~ concentrations (µM) are in rows for each strain.",
                    "Shapes show the imposed photoperiod (h); 4 h (hollow inverted triangle),  8 h (solid diamond), 12 h (solid circle), 16 h (solid upright triangle).",
                      "Symbol colours show the spectral waveband for growth; white LED (black symbols), 660 nm (red symbols), and 450 nm (blue symbols).",
                      "Large symbols show mean of growth rate from logistic curve fits; small symbols show values for replicate determinations, if any.",
                    "Platt (add Platt citation here) 4 parameter model fit to data pooled for each combination of strain and dissolved oxygen shown with solid lines. Seperate models fit to photoperiod data and shown if significantly different from the pooled model; 4 h (long dashed line); 8 h (dotted line); 12 h (dashed line); and 16 h (dot dashed line). ",
                       .sep = " ")

@@

MED4ProteinAnnotated_cap <- glue("**Bar graph indicating protein concentrations determined from lab grown *Prochlorococcus marinus* MED4 (High Light (HLI) near surface clade).**",

@@

MIT9313ProteinAnnotated_cap <- glue("**Bar graph indicating protein concentrations determined from lab grown *Prochlorococcus marinus* MIT9313 (Low Light (LLIV) deep ocean clade).  **",

>>>>

---
title: "*Prochlorococcus marinus* responses to light and oxygen" 
author:
  - name: '**Mireille Savoie**'
    affil: 1
    main: true
    email: msavoie@mta.ca
  - name: Aurora Mattison
    affil: 1, 2
  - name: Julie Nadeau
    affil: 1
  - name: Laurel Genge
    affil: 1
  - name: Sylwia Śliwińska-Wilczewska
    affil: 1
  - name: Maximilian Berthold
    affil: 1, 3
  - name: Naaman Omar
    affil: 1
  - name: Ondřej Prášil
    affil: 1, 4
    main: false
  - name: Amanda M. Cockshutt
    affil: 1, 5
    main: false
  - name: Douglas A. Campbell
    affil: 1
    main: false
    orcid: '0000-0001-8996-5463'
    email: 'dcampbel@mta.ca'
affiliation:
  - num: 1
    address: Department of Biology, Mount Allison University
institute:
  - mta: 'Department of Biology, Mount Allison University, 63B York St., Sackville NB, Canada, E4L 1G7.'
  - num: 2
  - ubc: 'Department of Biology, University of British Columbia'
  - num: 3
  - umwelt: 'Umweltbundesamtm, Dessau-Roßlau, Germany'
  - num: 4
  - alga: 'Algatech, Czech Academy of Science'
  - num: 5
  - sfx: 'Department of Chemistry, St. Frances Xavier University'
date: "`r format(Sys.Date())`"
output:
  bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
  bookdown::word_document2:
    code_folding: show
    reference_docx: KnitTemplate.docx
    keep_md: yes
    toc: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
always_allow_html: true
bibliography: [Manuscript_O2.bib, RPackages.bib]
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

@@

*Prochlorococcus marinus* comprises many strains, organized into clades, defined by 16S-23S intergenic transcribed ribosomal sequence signatures (Rocap et al., 2003). The clades inhabit distinct ecological niches [@moorePhysiologyMolecularPhylogeny1998], originally defined by Low-Light (LL) or High-Light (HL).  Of at least 12 known *Prochlorococcus* clades only 5 to date have cultured representatives; HLI, HLII, LLI, LLII/III and LIV [@billerProchlorococcusStructureFunction2015].

@@

Our changing climate, characterized by warming ocean waters, is rapidly altering conditions for these highly specialized strains of marine picophytoplankton. Predictions indicate a net global increase of *P. marinus* cell abundances of 29%, along with  poleward latitudinal shifts of at least 10° in marine phytoplankton niches by the end of this century [@flombaumPresentFutureGlobal2013] in response to warming waters, with increases in *Prochlorococcus* of approximately 50% in the more poleward regions. As the solar incidence angles are lower and photoperiod seasonality is more pronounced at higher latitudes, this shift will carry these phytoplankters into new combinations of photoperiod, seasonal regimes and changing spectral regimes [@bartonAnthropogenicClimateChange2016]. A 10° poleward shift of *P. marinus*, coupled with a dramatic increase in cell abundance between 40°N and 50°N,  could have a large impact on ocean ecosystems and global biogeochemical cycles [@flombaumPresentFutureGlobal2013], and thus will undoubtedly impact the ecosystem balance off Canadian waters.

Global warming is also rapidly changing the chemistry of our oceans.  By the end of this century, surface ocean pH is projected to decline by 0.1 to 0.4  due to  projected increases in carbon dioxide concentrations [@garcia-sotoOverviewOceanClimate2021]. Moreover, substantial changes in the global water cycle, leading to extensive changes in worldwide precipitation patterns, are affecting ocean salinity levels on a global scale, and  ice melts due to rising temperatures are impacting salinity levels in the Arctic and Northwest Atlantic oceans [@leeRecognizingSalinityThreats2022]. Increasing sea temperatures are also causing a decrease in [O~2~] across our global oceans [@matearLongtermChangesDissolved2003], with significant declines in [O~2~] observed toward the poles in both hemispheres [@helmObservedDecreasesOxygen2011]. Warmer ocean waters increase stratification, and decrease oxygen solubility at the surface, which in turn decreases oxygen mixing downwards by ocean currents [@garcia-sotoOverviewOceanClimate2021]. Models predict that the OMZ in the Pacific and Indian Oceans are expanding [@garcia-sotoOverviewOceanClimate2021; @buseckeDivergingFatesPacific2022]. In light of these expanding OMZ, a recent study by @buseckeDivergingFatesPacific2022 highlights an intriguing finding; the cores of the OMZ, where the oxygen levels are lowest, is projected to contract.  Given the predicted decrease in overall oceanic [O~2~], projected expansion of OMZs, and possible decline in the OMZ core volumes, we set forth to examine the influence of [O~2~] on *P. marinus* growth, as this area of research has been relatively understudied.

@@

We then analyzed growth rates in terms of cumulative diel Photosynthetically Usable Radiation (PUR). Photosynthetic organisms rely on absorbing light energy within the Photosynthetically Active Radiation (PAR) range, 350 to 700 nm, for photosynthesis [@morelAvailableUsableStored1978]. PUR represents the fraction of PAR that is absorbed by the pigments of photosynthetic organisms [@morelAvailableUsableStored1978]. PUR, therefore, takes into account the pigment composition of photosynthetic organisms and the specific spectral wavebands these pigments absorb.  *Prochlorococcus marinus* light-harvesting complexes, [@chisholmProchlorococcusMarinusNov1992;
@goerickePigmentsProchlorococcusMarinus1992], show an absorption maxima
of 442 nm for divinyl chlorophyll a and 478 nm for divinyl chlorophyll b
[@goerickePigmentsProchlorococcusMarinus1992] allowing *P. marinus* to
efficiently harvest blue light in the 400 nm
to 500 nm range [@morelAvailableUsableStored1978] prevailing in deep ocean
habitats, where only blue spectral wavelengths prevail [@holtropVibrationalModesWater2021].  In *Prochlorococcus* small cell diameters, from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992], and simple cell structures, minimize the complication of pigment package effect or intracellular self-shading [@morelProchlorococcusSynechococcusComparative1993]. Given the different spectral light regimes typical of the niches of different ecotypes,  expressing growth rates in terms of cumulative diel absorbed PUR might simplify different photoperiods, spectral bands, and PAR levels into a common parameter, making growth response comparisons across strains and different oxygen levels more accessible.  We also aimed to detect whether growth responses are driven simply by cumulative diel absorbed PUR, or whether specific photoperiods, spectral bands or PAR levels have independent, albeit interacting, effects on growth.

@@

Three xenic cultures of *P. marinus* were obtained from Bigelow Labs, NCMA
Maine, US (REFERENCE). MED4 (CCMP1986) is from High-Light adapted (HLI) clade; SS120
(CCMP1375) is from Low-Light adapted (LLIII) clade; and MIT9313 (CCMP2773)
is from Low-Light adapted (LLIV) clade. Cultures were
maintained in incubators set to 22°C with a light/dark cycle of 12 h. The PAR level for maintenance cultures reflected Photosynthetically Active Radiation (PAR) in the source
niche of the ecotype; MED4, of 160 µmol photons m^-2^ s^-1^ with illumination from STANDARD Products Inc. Cool White F24T5/41K/8/HO/PS/G5/STD, 24 watts, fluorescent bulbs; SS120 and
MIT9313 at 30 µmol photons m^-2^ s^-1^ with illumination from Philips Cool White F14T5/841 Alto, 14 watts, fluorescent bulbs. To maintain active growth all strains were
transferred weekly with 1 to 5 dilution with Pro99 media
[@mooreCulturingMarineCyanobacterium2007] prepared with autoclaved artificial
seawater [@bergesEvolutionArtificialSeawater2001].

@@

The growth rate of MED4, from the High-Light (HLI) clade of *P.
marinus*, increased with higher imposed PAR and longer photoperiods
(Figure \@ref(fig:MED4PAR)), when grown under 250 µM O~2~. No growth was
observed under any imposed conditions under a 4 h photoperiod. MED4 was
able to grow under the lowest PAR treatment of 30 µmol photons m^-2^
s^-1^ but only when the blue light photoperiod was longer than 4 h; the
white LED light photoperiod was longer than 8 h; or when the red light
photoperiod was longer than 16 h. Growth rate increased
with increasing photoperiods for all spectral wavebands when exposed to
PAR of 90 and 180 µmol photons m^-2^ s^-1^. MED4 grew fastest for each
spectral waveband under 16 h photoperiods. The maximum growth rate
(µ~max~) was
`r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 180 uE blue
light and 16 h photoperiod.

@@

MED4PAR_cap <- glue("**Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MED4 (High-Light (HLI) near surface clade) vs. photoperiod (h).  **",

@@

The growth rate of SS120, from the Low-Light (LLIII) clade of *P.
marinus*, increased with lower imposed PAR and longer photoperiods, when
grown under 250 µM O~2~ (Figure \@ref(fig:SS120PAR)). SS120 was able to
grow under the lowest PAR treatment of 30 µmol photons m^-2^ s^-1^ but
only when the photoperiod was longer than 4 h under blue, white, and red
light. SS120 required longer than 8 h photoperiod of imposed green
light. No growth was observed under any blue light photoperiods when
exposed to PAR of 90 µmol photons m^-2^ s^-1^. Growth rate increased
with increasing photoperiods for white, red and green light when exposed
to PAR of 90 µmol photons m^-2^ s^-1^; however, no growth was observed
when the red light photoperiod was 16 h. Experiments conducted under PAR
of 180 µmol photons m^-2^ s^-1^ showed no growth under imposed blue and
red lights. Growth rate decreased with longer photoperiods and showed
growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol
photons m^-2^ s^-1^ and white LED light. The growth rate generally
increased with longer photoperiods under green light treatments, however
no growth was observed under the 8 h photoperiod. The µ~max~ was
`r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^  achieved under 90 µmol photons
m^-2^ s^-1^ white LED light and 16 h photoperiod.

@@

SS120PAR_cap <- glue("**Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* SS120 (Low-Light (LLIII) deep ocean clade) vs. photoperiod (h). **",

@@

The growth rate of MIT9313, from the Low-Light (LLIV) clade of *P.
marinus*, increased with low to moderate imposed PAR and longer
photoperiods, when grown under 250 µM O~2~ (Figure
\@ref(fig:MIT9313PAR)). Growth rate increased with increasing
photoperiods under the lowest PAR treatment of 30 µmol photons m^-2^
s^-1^ for blue, white LED and red light. The green light treatment
showed growth only when the photoperiod was longer than 12 h and
decreased by \~ 50% under a 16 h photoperiod. Under PAR of 90 µmol
photons m^-2^ s^-1^, growth rate increased with longer photoperiods for
white LED, red, and green light treatments. Growth rates decreased from
4 to 8 h blue light photoperiod and exhibited no growth under
photoperiods of 12 and 16 h. Blue light did not induce growth at 180
µmol photons m^-2^ s^-1^ under any photoperiod; white LED and red light
treatments only exhibited growth under the 8 h photoperiod and growth
rates increased with longer photoperiods under green light. The µ~max~
was
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 30 µmol photons
m^-2^ s^-1^ blue light and 16 h photoperiod.

@@

MIT9313PAR_cap <- glue("**Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MIT9313 (Low-Light (LLIV) deep ocean clade) vs. photoperiod (h). **",

@@

AllStrainPurFitsPlots_cap <- glue("**Chlorophyll specific growth rate (d^-1^) vs. cumulative diel Photosynthetic Usable Radiation (PUR, µmol photons m^-2^ d^-1^).**",
                    "**A.** MED4 High-Light I ecotype under 250, 25, or 2.5 µM of O~2~.",
                    "**B.** SS120 Low-Light II/III (LLII/III) ecotype under 250, 25, or 2.5 µM of O~2~.",
                    "**C.** MIT9313 Low-Light IV (LLIV) ecotype under 250, 25, or 2.5 µM of O~2~.",
                    "3 levels of imposed dissolved O~2~ concentrations (µM) are in rows for each strain.",
                    "Shapes show the imposed photoperiod (h); 4 h (hollow inverted triangle),  8 h (solid diamond), 12 h (solid circle), 16 h (solid upright triangle).",
                      "Symbol colours show the spectral waveband for growth; white LED (black symbols), 660 nm (red symbols), and 450 nm (blue symbols).",
                      "Large symbols show mean of growth rate from logistic curve fits; small symbols show values for replicate determinations, if any.",
                    "Platt (add Platt citation here) 4 parameter model fit to data pooled for each combination of strain and dissolved oxygen shown with solid lines. Seperate models fit to photoperiod data and shown if significantly different from the pooled model; 4 h (long dashed line); 8 h (dotted line); 12 h (dashed line); and 16 h (dot dashed line). ",
                       .sep = " ")

@@

MED4ProteinAnnotated_cap <- glue("**Bar graph indicating protein concentrations determined from lab grown *Prochlorococcus marinus* MED4 (High-Light (HLI) near surface clade).**",

@@

MIT9313ProteinAnnotated_cap <- glue("**Bar graph indicating protein concentrations determined from lab grown *Prochlorococcus marinus* MIT9313 (Low-Light (LLIV) deep ocean clade).  **",
